The watershed transform in ITK - discussion and new developments

### 2006-05-17 article/Article.tex
marker watershed timing
Added timing for marker based watershed. Not sure why there was such a big
difference between fully and face connected for the itk watershed. Difference
due to minima imposition isn't enough to account for it.

darcs-hash:20060517040544-fafb9-ef02a7f12e50e5a2ea52bd3a54b05b93237644fd.gz
continue the article
darcs-hash:20060517134412-2fc9d-c1dc3633a52f22ceb1f8894fb8ea63a10853ec4c.gz

@@

This can be achieved by performing a {\em reconstruction by erosion}
or {\em recursive conditional erosion} from the seleted regional
minima. Figure \ref{fig:recon} illustrates the results of apply a
reconstruction by erosion to a 1D image. The original image is shown
in blue and has regional minima A, B, C, D and E. The marker image
which is recursively eroded is shown in black, with a single zero at
regional minimum B. The red plateaus illustrate the parts of the
result image that are above the orginal. The resulting image has only
on regional minima that corresponds to B.
\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.5]{recon}
\caption{Reconstruction by erosion from a single regional minimum.}
\label{fig:recon}
\end{center}
\end{figure}

@@

\begin{table}[htbp]
\centering
\begin{tabular}{cccccccc}
\hline
FC & MB & Find Min &  Label mins & Watershed from markers & Total & Morph watershed &  itk watershed \\
\hline
0  &    0  &     1.948  & 0.522  & 5.752  & 8.222   & 8.218  & 12.596 \\
0  &    1  &     1.922  & 0.516  & 6.724  & 9.162   & 9.28   & 12.61 \\
1  &    0  &     4.26   & 0.154  & 13.274 & 17.688  & 18.158 & 12.61 \\
1  &    1  &     4.284  & 0.15   & 17.23  & 21.664  & 21.76  & 12.606 \\
\hline
\end{tabular}
\caption{Execution times. Note the FC and MB refer to FullyConnected and MarkBoundary.\label{tbl:perf}}
\end{table}

  \subsection{Marker based watershed}
\label{sect:perf-marker}

The performance in case of a marker based approach are shown in Table
\ref{tbl:perf-marker}, and were obtained on the same system than the
ones in Section \ref{sect:perf-classical}, with the command {\tt 
./mperf3D images/ESCells.img images/ESCells-markers.tif}.
The ITK watershed use the implementation presented in Section
\ref{sect:implementation}. The extra steps of this implementation are
also shown in this table.

\begin{table}[htbp]
\centering
\begin{tabular}{cccccccc}
\hline
FC & MB & impose min &  itk watershed & match labels & change labels & total & Morph watershed \\
\hline
\hline
\end{tabular}
\caption{Execution times with marker based approach. Note the FC and MB refer to FullyConnected and MarkBoundary.\label{tbl:perf-marker}}
\end{table}

@@

\section{Using watersheds in ITK}

@@

The program is in python for conciseness\footnote{And also to promote
python wrappers.}.

@@

\small \begin{verbatim}
ws = itk.MorphologicalWatershedImageFilter.IUC2IUC2.New(idistance,
                                                        MarkWatershed=False,
                                                        Level=10)
mask = itk.MaskImageFilter.IUC2IUC2IUC2.New(ws, fill)
overlay = itk.LabelOverlayImageFilter.IUC2IUC2IRGBUC2.New(reader, mask)
\end{verbatim} \normalsize

@@

\small \begin{verbatim}
ws2 = itk.WatershedImageFilter.IUC2.New(idistance, Level=0.5)
relabel = itk.RelabelComponentImageFilter.IUL2IUC2.New(ws2)
mask2 = itk.MaskImageFilter.IUC2IUC2IUC2.New(relabel, fill)
overlay2 = itk.LabelOverlayImageFilter.IUC2IUC2IRGBUC2.New(reader, mask2)
\end{verbatim} \normalsize

ITK watershed clearly produce a wrong segmentation in that case.
The result of the morphological watershed is quite good, even if not perfect:
there are some curves in the separation line of the two labels. This is the
result of the distance approximation in the watershed algorithm.

>>>>

Section \ref{sect:using-ws-itk} propose some applications of watershed
with ITK.

@@

This can be achieved by performing a {\em reconstruction by erosion}
or {\em recursive conditional erosion} from the seleted regional
minima. This process is also called {\em minima imposition}. Figure
\ref{fig:recon} illustrates the results of apply a
reconstruction by erosion to a 1D image. The original image is shown
in blue and has regional minima A, B, C, D and E. The marker image
which is recursively eroded is shown in black, with a single zero at
regional minimum B. The red plateaus illustrate the parts of the
result image that are above the orginal. The resulting image has only
on regional minima that corresponds to B.
\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.5]{recon}
\caption{Reconstruction by erosion from a single regional minimum.}
\label{fig:recon}
\end{center}
\end{figure}

@@

\begin{table}[htbp]
\centering
\begin{tabular}{cc|ccc|cc}
\hline
FC & MB & Find min &  Label min & Ws from markers & Morph ws &  ITK ws \\
\hline
\hline
0 &      0 &      1.61 &   0.0476 & 3.72 &   6.14 &   - \\
0 &      1 &      1.59 &   0.0428 & 6.25 &   8.95 &   - \\
1 &      0 &      3.15 &   0.0465 & 7.77 &   11.5 &   6.48 \\
1 &      1 &      3.19  &  0.0462  &14.8 &   19.2  &  - \\
% 0  &    0  &     1.948  & 0.522  & 5.752  & 8.222   & 8.218  & 12.596 \\
% 0  &    1  &     1.922  & 0.516  & 6.724  & 9.162   & 9.28   & 12.61 \\
% 1  &    0  &     4.26   & 0.154  & 13.274 & 17.688  & 18.158 & 12.61 \\
% 1  &    1  &     4.284  & 0.15   & 17.23  & 21.664  & 21.76  & 12.606 \\
\hline
\end{tabular}
\caption{Execution times. Note the FC and MWL refer to FullyConnected and MarkWatershedLine.\label{tbl:perf}}
\end{table}

\subsection{Marker based watershed}
\label{sect:perf-marker}
The performance in case of a marker based approach are shown in Table
\ref{tbl:perf-marker}, and were obtained on the same system than the
ones in Section \ref{sect:perf-classical}, with the command {\tt 
./mperf3D images/ESCells.img images/ESCells-markers.tif}.
The ITK watershed use the implementation presented in Section
\ref{sect:implementation}. The extra steps of this implementation are
also shown in this table.

\begin{table}[htbp]
\centering
\begin{tabular}{cc|cccc|cc}
\hline
FC & MWL & Impose min &  ITK ws & Match labels & Change labels & Morph ws & ITK ws \\
\hline
\hline
0 &      0 &      -    &   -   &    -      & -     &  3.67 &   - \\
0 &      1 &      -    &   -   &    -      & -     &  6.53 &   - \\
1 &      0 &      5.46 &   7.8 &    0.0192 & 0.227 &  14.6 &   13.5 \\
1 &      1 &      -    &   -   &    -      & -     &  34.8 &   - \\
% 0   &  0  &  2.47776 & 8.12256 & 4.29153e-07 & 0.0990558 &  10.6994 & 3.70646\\
% 0   &  1  &  2.24303 & 8.20924 & 9.53674e-07 & 0.0977524 &  10.55   & 6.22279\\
% 1   &  0  &  5.92089 & 12.2889 & 4.29153e-07 & 0.0971172 &  18.3069 & 10.7172\\
% 1   &  1  &  6.08657 & 12.222  & 5.72205e-07 & 0.0927572 &  18.4014 & 18.4605\\
\hline
\end{tabular}
\caption{Execution times with marker based approach. Note the FC and MWL refer to FullyConnected and MarkWatershedLine.\label{tbl:perf-marker}}
\end{table}

@@

\section{Using watersheds in ITK}
\label{sect:using-ws-itk}

It should be noticed that all the images shown in this article have been
produced with ITK and can be easily reproduced and tested by running 
{\em ctest} in the build tree.

@@

The program is in python for conciseness\footnote{And also to promote
python wrappers.}. A C++ version, \verb$splitPronucleus2D.cxx$, is also
available.

@@

\small \begin{verbatim}
ws = itk.MorphologicalWatershedImageFilter.IUC2IUC2.New(idistance,
                                                        MarkWatershedLine=False,
                                                        Level=10)
mask = itk.MaskImageFilter.IUC2IUC2IUC2.New(ws, fill)
overlay = itk.LabelOverlayImageFilter.IUC2IUC2IRGBUC2.New(reader,
                                                          mask,
                                                          UseBackground=True)
\end{verbatim} \normalsize

@@

\small \begin{verbatim}
ws2 = itk.WatershedImageFilter.IUC2.New(idistance, Level=0.5)
relabel = itk.RelabelComponentImageFilter.IUL2IUC2.New(ws2)
mask2 = itk.MaskImageFilter.IUC2IUC2IUC2.New(relabel, fill)
overlay2 = itk.LabelOverlayImageFilter.IUC2IUC2IRGBUC2.New(reader,
                                                           mask2
                                                           UseBackground=True)
\end{verbatim} \normalsize

ITK watershed clearly produce a wrong segmentation in that case. However, using
a floating point image to store the distance map would have produce a better
result.
The segmentation of the morphological watershed is quite good, even if not perfect:
there are some curves in the separation line of the two labels. This is the
result of the distance approximation in the watershed algorithm.
